---
title: "Codebook"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
This Codebook refers to the dataset “Building registry of Vienna in the late 1920s”, briefly called digital building registry. The digital building registry is the machine-readable version of [Häuser-Kataster der Bundeshauptstadt Wien](https://permalink.obvsg.at/wbr/AC07637508). The transformation from the analog to the digital version is documented by a Data Descriptor, published in Nature Scientific Data.

##	Dataset format
The dataset has a CSV format. The data records are separated by semicolons (“;”). The decimal separator is a decimal point “.”. The Universal Coded Character Set (UCS) is UTF-8. 

##	Dataset fields
The dataset includes 15 fields, as listed below. It includes fields from the analog building registry, briefly called original data (OD) and supplementary data (SD), which have been added to improve data processing and validation as well as usability. It is noted that the analog building registry consists of ten volumes, which have been published between 1927 and 1930 (online table: codebook_analog.building.registry.csv). So, we attached a “.1920s” to these database fields. The “.2010s” addition indicates that the timestamp is in the late 2010s – between 2018 and 2019.

## Comments
This section comments on the individual data fields by giving background information and details for using the data.

###	ID
It is noted that the analog building registry has entries with multiple building numbers per entry. These entries have been separated to have only a single building number per data entry.
The field “ID” is a character that stands for a unique address. An address is the combination of the street name and the building number.
The next figure exemplifies the address relation between the analog and the digital building registry.

[test](Codebook_files/table_codebook.table.csv)


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
