---
title: "Codebook"
output:
  github_document:
    toc: true
    toc_depth: 3


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openxlsx)
library(rmarkdown)
library(knitr)


# Import dataset
dataset <- read.table("C:/Users/u.kral/ownCloud/03_TU Wien/Github/building.schematic/Dataset.csv", sep = ";", header = TRUE, encoding="UTF-8", stringsAsFactors = F)


# Import tables and figures
table1 <- read.table("table_dataset.datafields.csv", sep = ";", header = TRUE, encoding="UTF-8")
colnames(table1) <- c("Field", "Name", "Name.analog", "Description", "Type")

table_id <- read.table("table_ID.csv", sep = ";", header = TRUE, encoding="UTF-8")
table_STR.2010s <- read.table("table_STR.2010s.csv", sep = ";", header = TRUE, encoding="UTF-8")
table_POS.1920s <- read.table("table_POS.1920s.csv", sep = ";", header = TRUE, encoding="UTF-8")
table_FLOORS.1920s <- read.table("table_FLOORS.1920s.csv", sep = ";", header = TRUE, encoding="UTF-8")
```


# Introduction
This Codebook refers to the dataset [Building schematic of Vienna in the late 1920s](https://github.com/ukral/building.schematic/blob/master/Dataset.csv), briefly called digital building schematic. The dataset includes `r nrow(dataset)` data entries (rows) and `r ncol(dataset)` data fields (columns).

#	Dataset format
The dataset has a CSV format. The data records are separated by semicolons (“;”). The decimal separator is a decimal point “.”. The Universal Coded Character Set (UCS) is UTF-8. 

#	Dataset fields
The dataset includes `r nrow(table1)` data fields. Eight data fields originate from the analog building schematic, briefly called original data (OD) and four data fields that have been supplemented (SD). The suffixes “.1920s” and “.2010s” indicate data records in the late 1920s and late 2010s, respectively.

```{r, echo=FALSE}
kable(head(table1,nrow(table1)))

```

# Comments
This section comments on the individual data fields by giving background information and details for using the data.

##	ID
It is noted that the analog building schematic has entries with multiple building numbers per entry. These entries have been separated to have only a single building number per data entry.

The field “ID” is a character that stands for a unique address. An address is the combination of the street name and the building number.
The next figure exemplifies the address relation between the analog and the digital building schematic. For instance, the address "lerchenfelderstraße 100,102" is converted into the following data entries:

```{r, echo=FALSE}
kable(head(table_id))

```

## UD.1920s
The field specifies the urban district in which the building is located in the late 1920s. It is noted that the district boundaries differ from those in the late 2010s. It is also stated that in the late 1920s the city covered only the districts 1 to 21. The districts 22 and 23 were amalgamated with the city after the 1920s.

## CC.2010s
The field specifies a numerical code of the cadastral community as given in the late 2010s. It is noted that the geospatial data on administrative boundaries of the city can be retrieved from the Austrian [open government data plattform](https://www.data.gv.at/katalog/dataset/verwaltungsgrenzen-vgd-stichtagsdaten-wien).

##	STR.1920s
This field specifies the name of the street, alley or square as given in the late 1920s.

## STR.2010s
This field specifies two type of entries.
```{r, echo=FALSE}
kable(head(table_STR.2010s))
```

##	BN.1920s
The field specifies the number of a building in a street as given in late 1920s. An empty cell indicates no data record.


##	POS.1920s
The field specifies the position of the building within a row of buildings.
```{r, echo=FALSE}
kable(head(table_POS.1920s))
```

##	AREA.1920s
The field specifies the property area in m2. “NA” stands for “not available”.

##	FLOORS.1920s
The meaning of blank records is ambiguous and is considered to be unknown unless additional efforts are undertaken to replace them by evidence-based data from historical construction plans. The integer floor numbers represent the floors above a ground floor and mezzanine with a lower base floor, respectively. So, a building with a data record "1" has two floors in total, one ground floor and one above.
```{r, echo=FALSE}
kable(head(table_FLOORS.1920s))
```

##	YoC.1920s
The field covers the year in which the building was constructed. Blank entries stand for “not available”.

##	YoP.1920s
The data field covers the year the building was transferred to the new owner. Blank entries stand for “not available”.

##	Page.pdf
The field includes integers that stand for the page number in the scanned PDF version of the [analog building schematic](https://permalink.obvsg.at/wbr/AC07637508).

# Exemplary data records
```{r, echo=FALSE}
kable(head(dataset[1:5,]))
```


