---
title: "Technical Validation"
author: "Ulrich Kral"
date: "14 10 2020"
output:
  html_document:
    
    code_folding: hide
    fig_caption: true
    number_sections: true
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())

library(plyr)
library(dplyr)
library(naniar)
library(readxl)
library(ggplot2)
library(patchwork)
library(splus2R)
library(ggpubr)

```

# Preface
This code corresponds with the Technical Validation section in the Data Descriptor "Building schematic of Vienna in the late 1920s", published by Nature Scientific Data.

Please consider the following steps to run the code
1. Create a new directory on your computer (e.g. "c:/building.schematic")
2. Download the files from the [Github repository](https://github.com/ukral/building.registry) into your new directory.
3. Copy the path of your new directory into the code at line 42.

# Import datasets

```{r}
######################################################################################
path <- "C:/Users/u.kral/ownCloud/03_TU Wien/Github/building.schematic/"
######################################################################################


# Import file "Dataset.csv", which is the digital building schematic.
dataset <- read.csv(file=paste(path, "Dataset.csv", sep = ""), sep = ";", stringsAsFactors = FALSE, encoding = "UTF-8")

# Import file "Online-Only Table 2.csv", which is identical with Online-only Table 2 in the Data Descriptor.
cadastral_raw <- read.csv(file=paste(path, "/Data.descriptor/Online-only Table 2.csv", sep=""),  sep = ";" , stringsAsFactors = F)
cadastral <- cadastral_raw[1:66,]

# Import file "adressen_standorte_wien_20201015.csv". This file includes today's street names. The is available on the Open Government Data platform and can be retrieved from https://www.data.gv.at/katalog/dataset/stadt-wien_adressdatenderstadtwien. We retrieved the file at 15 October 2020 and copied it the the Github repository.
adressen <- read.csv(file=paste(path, "/Data.descriptor/adressen_standorte_wien_20201015.csv", sep=""), sep = ";", stringsAsFactors = F,fileEncoding = "UTF-8")

# Import file "statistical_yearbook (1914).xlsx"
floors_1914 <- data.frame(read_xlsx(paste(path, "/Data.descriptor/statistical_yearbook (1914).xlsx", sep=""), sheet = "STKW_hist", col_names = F, range = "B7:I27"), stringsAsFactors = F)

colnames(floors_1914) <- c("UD.1920s", "FLOORS_0", "FLOORS_1", "FLOORS_2", "FLOORS_3", "FLOORS_4", "FLOORS_5", "FLOORS_unknown")

# Import file "statistical_yearbook (1923).xlsx"
yearbook_1923 <- read_xlsx(paste(path, "/Data.descriptor/statistical_yearbook (1923).xlsx", sep=""), sheet = "Rohdaten", col_names =  TRUE, col_types = rep("numeric", times = 2))




```


# Internal validation
This code section produces Figure 6 in the Data Descriptor.
## ID
```{r}

id_0 <- grep("_[0-9]*", dataset$ID) # aufgefalteteten IDs
id_1 <- grep("^[0-9]*$", dataset$ID) # nicht aufgefaltet
id_2 <- grep("_[1]$", dataset$ID) # nur die mit _1

id_plot_data <- dataset
id_plot_data$id_flag <- rep(NA, nrow(dataset))
id_plot_data[id_0, "id_flag"] <- "yes" # logic: ID unfolded: yes
id_plot_data[id_1, "id_flag"] <- "no"

id_plot_data_grouped <- id_plot_data %>%
  group_by(UD.1920s, id_flag) %>%
  summarize(count = n())

id_plot_data_grouped <- data.frame(id_plot_data_grouped, stringsAsFactors = F)

totals<- id_plot_data_grouped %>%
	group_by(UD.1920s) %>%
	summarise(total = sum(count))

# Plot
id_plot <- ggplot(id_plot_data_grouped, aes(x=UD.1920s, y=count, fill= factor(id_flag, levels = c("yes", "no")))) +
  geom_bar(stat="identity", position="stack")+
  labs(title = "ID", x="Urban district (UD.1920s)", y="Number of data entries")+
  theme(plot.title = element_text(face = "bold"))+
  scale_fill_manual(name = "ID unfolded", values=c('lightblue','darkblue'))+
  geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150, size = 3)+
  scale_x_continuous(breaks = c(1:21))
id_plot


```

## STR.1920s
```{r}


str_ma37 <- as.character(levels(as.factor(adressen$NAME_STR)))

# Standardizing the street names in "Adressen Standorte Wien", which is the external datasource for validating STR.1920s and STR.2010s names in the dataset
str_m <- gsub(" ","", str_ma37) 
str_m <- lowerCase(str_m) 
str_m <- gsub("st\\.", "sankt", str_m)
str_m <- gsub("-","",str_m) 
str_m <- gsub("dr\\.","dr",str_m) 
str_m <- gsub("ß","ss",str_m)
str_m <- gsub("'","",str_m)
str_m <- gsub("\\.","",str_m)

str_gesamt <- data.frame(str_m, str_m, stringsAsFactors = FALSE)
colnames(str_gesamt) <- c("str_gesamt", "str_ma37")

# Standardizing the STR.1920s names in the dataset
dataset$str_1920_norm <- dataset$STR.1920s
dataset$str_1920_norm <- lowerCase(dataset$str_1920_norm) 
dataset$str_1920_norm <- gsub("ß","ss", dataset$str_1920_norm)
dataset$str_1920_norm <- gsub("st\\.","sankt",dataset$str_1920_norm)
dataset$str_1920_norm <- gsub("dr\\.","dr",dataset$str_1920_norm)
dataset$str_1920_norm <- gsub(" ","", dataset$str_1920_norm) 
dataset$str_1920_norm <- gsub("'","",dataset$str_1920_norm)
dataset$str_1920_norm <- gsub("\\.","", dataset$str_1920_norm)
dataset$str_1920_norm <- gsub("-","", dataset$str_1920_norm)
dataset$str_1920_norm <- gsub("\\(","", dataset$str_1920_norm)
dataset$str_1920_norm <- gsub("\\)","", dataset$str_1920_norm)

dataset$str_2010_norm <- dataset$STR.2010s
dataset$str_2010_norm <- lowerCase(dataset$str_2010_norm) # alles auf Kleinbuchstaben
dataset$str_2010_norm <- gsub("ß","ss", dataset$str_2010_norm)
dataset$str_2010_norm <- gsub("st\\.","sankt",dataset$str_2010_norm)
dataset$str_2010_norm <- gsub("dr\\.","dr",dataset$str_2010_norm)
dataset$str_2010_norm <- gsub(" ","", dataset$str_2010_norm) # Leerzeichen entfernen
dataset$str_2010_norm <- gsub("'","",dataset$str_2010_norm)
dataset$str_2010_norm <- gsub("\\.","", dataset$str_2010_norm)
dataset$str_2010_norm <- gsub("-","", dataset$str_2010_norm)
dataset$str_2010_norm <- gsub("\\(","", dataset$str_2010_norm)
dataset$str_2010_norm <- gsub("\\)","", dataset$str_2010_norm)

# Assigning standardized names from "Adressen Standorte Wien" to the standardized names of STR.2010s.
dataset <- merge(dataset, str_gesamt, by.x = "str_2010_norm", by.y = "str_gesamt", all.x = TRUE)

# Assigning standardized names from "Adressen Standorte Wien" to the standardized names of STR.1920s.
dataset <- merge(dataset, str_gesamt, by.x = "str_1920_norm", by.y = "str_gesamt", all.x = TRUE)

length(levels(as.factor(dataset$str_ma37.y))) 

match_yes<- dataset[which(dataset$str_ma37.y != ""),] 
match_no <- dataset[which(is.na(dataset$str_ma37.y) == T),] 

match_yes_control <- dataset[which(dataset$str_1920_norm == dataset$str_2010_norm), ] 
match_no_control <- dataset[which(dataset$str_1920_norm != dataset$str_2010_norm), ] 

test1 <- match_no_control[which(is.na(match_no_control$str_ma37.y) == F & is.na(match_no_control$str_ma37.x) == F),] # 580 Straßenname hat sich geändert.
test2 <- match_no_control[which(is.na(match_no_control$str_ma37.y) == F & is.na(match_no_control$str_ma37.x) == T),] # 129, keine entsprechung bzw. aus dem Verkehrsflächenverzeichnis gestrichen.
test3 <- match_no_control[which(is.na(match_no_control$str_ma37.y) == T &is.na(match_no_control$str_ma37.x) == F),] # 1405, Straßenname hat sich geändert.
test4 <- match_no_control[which(is.na(match_no_control$str_ma37.y) == T &is.na(match_no_control$str_ma37.x) == T),] # 258, keine entsprechung bzw. aus dem Verkehrsflächenverzeichnis gestrichen.


# unique records
test1_uni <- unique(test1[,c("str_1920_norm","str_2010_norm")])
test2_uni <- unique(test2[,c("str_1920_norm","str_2010_norm")])
test3_uni <- unique(test3[,c("str_1920_norm","str_2010_norm")])
test4_uni <- unique(test4[,c("str_1920_norm","str_2010_norm")])

# Numbers for the data descriptor
str_1920_spelling_rows <- cbind(c(nrow(match_yes_control), nrow(match_no_control ), nrow(test1), nrow(test2), nrow(test3), nrow(test4)), round(c(nrow(match_yes_control), nrow(match_no_control ), nrow(test1), nrow(test2), nrow(test3), nrow(test4))/nrow(dataset)*100,2))

dataset$str_1920_spelling = c(rep(NA, nrow(dataset)))

# Data preperation for barplots

pos_str_1920_spelling_1 <- c(which(dataset$str_1920_norm == dataset$str_2010_norm), which(is.na(match_no_control$str_ma37.y) == F & is.na(match_no_control$str_ma37.x) == F), which(is.na(match_no_control$str_ma37.y) == F & is.na(match_no_control$str_ma37.x) == T))

pos_str_1920_spelling_2 <- c(which(is.na(match_no_control$str_ma37.y) == T &is.na(match_no_control$str_ma37.x) == F),which(is.na(match_no_control$str_ma37.y) == T &is.na(match_no_control$str_ma37.x) == T))

length(pos_str_1920_spelling_1)+length(pos_str_1920_spelling_2)-nrow(dataset)

str_1920_spelling_t1 <- c(match_yes_control[,"ID"], test1[,"ID"], test2[,"ID"])
str_1920_spelling_t2 <- c(test3[,"ID"], test4[,"ID"])


dataset[(dataset$ID %in% str_1920_spelling_t1), "str_1920_spelling"] <- "1" # Adressen Standorte Wien
dataset[(dataset$ID %in% str_1920_spelling_t2), "str_1920_spelling"] <- "2" # # Wien Geschichte Wiki, Analog Building Schematic


rows_pos_str_1920_spelling <- dataset %>%
  group_by(UD.1920s, str_1920_spelling) %>%
  summarize(count_spelling_1920 = n())


################

dataplot <- data.frame(rows_pos_str_1920_spelling, stringsAsFactors = F)
dataplot <- rbind(dataplot, c(8,2,0))
dataplot <- dataplot[order(dataplot$UD.1920s, dataplot$str_1920_spelling),]

totals<- dataplot %>%
	group_by(UD.1920s) %>%
	summarise(total=sum(count_spelling_1920))


# Plot
str_1920_plot <- ggplot(dataplot, aes(x=UD.1920s, y=count_spelling_1920, fill= factor(str_1920_spelling, levels = c("2","1")))) +
  geom_bar(stat="identity", position="stack") +
  labs(title = "STR.1920s", x="Urban district (UD.1920s)", y="Number of data entries") +
  scale_fill_manual(name = "Name spelling\nverified by", labels = c("Digital building schematic,\nWien Geschichte Wiki", "Adressen Standorte Wien"), values=c('lightblue','darkblue')) +
  geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150, size = 3) +
  scale_x_continuous(breaks = c(1:21)) +
  theme(plot.title = element_text(face = "bold"))
str_1920_plot

count_spelling_1920_sum <- nrow(dataset) - sum(dataplot$count_spelling_1920)



```


## STR.2010s
```{r}

str_2010_1 <- dataset[which(is.na(dataset$str_ma37.x) == FALSE),] 
str_2010_2 <- dataset[which(is.na(dataset$str_ma37.x) == TRUE),]

str_2010_sum <- nrow(dataset) - length(str_2010_1) - length(str_2010_2)

str_2010_spelling_t1 <- str_2010_1[,"ID"]
str_2010_spelling_t2 <- str_2010_2[,"ID"]

dataset[(dataset$ID %in% str_2010_spelling_t1), "str_2010_spelling"] <- "1" # Adressen Standorte Wien
dataset[(dataset$ID %in% str_2010_spelling_t2), "str_2010_spelling"] <- "2" # keine STR.2010s

rows_pos_str_2010_spelling <- dataset %>%
  group_by(UD.1920s, str_2010_spelling) %>%
  summarize(count_spelling_2010 = n())


dataplot <- data.frame(rows_pos_str_2010_spelling, stringsAsFactors = F)
dataplot <- rbind(dataplot, c(8,2,0))
dataplot <- dataplot[order(dataplot$UD.1920s, dataplot$str_2010_spelling),]

totals<- dataplot %>%
	group_by(UD.1920s) %>%
	summarise(total=sum(count_spelling_2010))


# Plot
str_2010_plot <- ggplot(dataplot, aes(x=UD.1920s, y=count_spelling_2010, fill= factor(str_2010_spelling, levels = c("2","1")))) +
  geom_bar(stat="identity", position="stack") +
  labs(title = "STR.2010s", x="Urban district (UD.1920s)", y="Number of data entries") +
  scale_fill_manual(name = "Name spelling\nverified by", labels = c("not relevant, because no STR.2010s\ncounterpart from STR.1920s", "Adressen Standorte Wien"), values=c('#999999','darkblue')) +
  geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150, size = 3) +
  scale_x_continuous(breaks = c(1:21))+
  theme(plot.title = element_text(face = "bold"))
str_2010_plot


count_spelling_2010_sum <- nrow(dataset) - sum(dataplot$count_spelling_2010)


```


## UD.1920
```{r}

ud <- dataset %>%
  group_by(UD.1920s) %>%
  summarize(count = n())
ud <- data.frame(ud, stringsAsFactors = F)

ud_merge <- merge(ud, unique(cadastral[,c("UD.1920s", "Volume")]), by.x = "UD.1920s", by.y = "UD.1920s", sort = F)
ud_merge <- data.frame(ud_merge, stringsAsFactors = F)

ud_plot <-ggplot(data=ud_merge, aes(x=UD.1920s, y=count, fill = factor(Volume, levels = c(1:10)))) +
  geom_bar(stat="identity", width=.8) +
  geom_text(aes(label=count), nudge_y=120, size = 3)+
  labs(title = "UD.1920s", x="Urban district (UD.1920s)", y="Number of data entries") +
  scale_x_continuous(breaks = c(1:21))+
  theme(plot.title = element_text(face = "bold"))+
  scale_fill_discrete("Volume of analog\nbuilding schematic")
ud_plot

ud_sum <- nrow(dataset) - sum(ud$count)
```


## CC.2010s

```{r}

cc <- dataset %>%
  group_by(CC.2010s) %>%
  summarize(count = n())

cc <- data.frame(cc, stringsAsFactors = F)
cc$CC.2010s <- as.character(cc$CC.2010s)

cadastral_raw_filter <- cadastral_raw[grep("\\d", cadastral_raw$cadastral.number_1920s),]
cadastral_raw_filter$cadastral.number_2010s <- as.character(cadastral_raw_filter$cadastral.number_2010s)

cc_merge <- merge(cc, cadastral_raw_filter, by.x = "CC.2010s", by.y = "cadastral.number_2010s", by=all, sort = F)
cc_merge  <- data.frame(cc_merge , stringsAsFactors = F)

totals<- cc  %>%
	group_by(CC.2010s) %>%
	summarise(total=sum(count))

cc_plot <-ggplot(data=cc_merge, aes(x=CC.2010s, y=count, fill = factor(Volume, levels = c(1:10)))) +
  geom_bar(stat="identity") +
  geom_text(aes(label=count), vjust=0.3, hjust=-.5, size=2.5, angle = 90) +
  labs(title = "CC.2010s", x="Cadastral communites (CC.2010s)",y="Number of data entries") +
  theme(axis.text.x=element_text(angle = 90, size = 6),plot.title = element_text(face = "bold"))+
  coord_cartesian(ylim = c(0,round(max(cc$count),-3)))+
  scale_fill_discrete(name = "Volume of analog\nbuilding schematic")
cc_plot

cc_sum <- nrow(dataset) - sum(cc_merge$count)
```

## BN.1920s

```{r}

# Define categories

pos1 <- grep("\\D$", dataset$BN.1920s) # 
test1 <- levels(as.factor(dataset[pos1, "BN.1920s"])) # 118b

pos2 <- grep("\\d$", dataset$BN.1920s) # 
test2 <- levels(as.factor(dataset[pos2, "BN.1920s"])) # 118 inkl. 2 Einträge für "neben27"

pos3 <- which(dataset$BN.1920s =="")
test3 <- dataset[pos3,]

pos4 <- grep("^[A-Za-z]", dataset$BN.1920s)
test4 <- levels(as.factor(dataset[pos4, "BN.1920s"])) # 2 Einträge für "neben27"

# die "neben27" rausfiltern
t <- which(pos2 %in% pos4)
pos2 <- pos2[-t] 

pos_test <- grep("^[1-9]*$", dataset$BN.1920s) # 
test <- levels(as.factor(dataset[pos_test, "BN.1920s"])) # Nur Zahlen inkl blanks. Bekomme die Blanks nicht weg


pos_comb <- c(pos1, pos2, pos3)
pos_c <- length(pos1)+length(pos2)+length(pos3)

bn.count <- c(length(pos2)-2, length(pos1), length(pos4), length(pos3))
bn.discr <- c("Only Integer", "Integer and letters", "Letters and integers", "Data not available")

bn.table <- cbind(bn.discr, bn.count)
bn.table <- as.data.frame(bn.table, stringsAsFactors = F)
bn.table$bn.count <- as.numeric(bn.table$bn.count)
bn.table$rel <- bn.table$bn.count / sum(bn.table$bn.count)*100

dataset$bn.table.ud = rep(NA, nrow(dataset))
dataset[pos1,"bn.table.ud"] <- bn.discr[2]
dataset[pos2,"bn.table.ud"] <- bn.discr[1]
dataset[pos3,"bn.table.ud"] <- bn.discr[4]
dataset[pos4,"bn.table.ud"] <- bn.discr[3]

####


bn.table_ud <- dataset %>%
  group_by(UD.1920s, bn.table.ud) %>%
  summarize(count = n())

test_sum <- nrow(dataset)- sum(bn.table_ud$count)

# stacked barplot
bn.table_ud <- data.frame(bn.table_ud, stringsAsFactors = F)

totals<- bn.table_ud %>%
	group_by(UD.1920s) %>%
	summarize(total=sum(count))

position_bn <- levels(as.factor(bn.table_ud$bn.table.ud))

bn_plot <-ggplot(data=bn.table_ud, aes(x=UD.1920s, y=count, fill = factor(bn.table.ud, levels = position_bn[c(1,3,2,4)]))) +
  geom_bar(stat="identity") +
  scale_fill_manual(name = "Data pattern", values=c('#999999','#87CEFA','#4169E1','darkblue')) +
  labs(title = "BN.1920s", x="Urban district (UD.1920s)", y="Number of data entries") +
  geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150, size = 3)+
  scale_x_continuous(breaks = c(1:21))+
  theme(plot.title = element_text(face = "bold"))
bn_plot


```

## AREA.1920s

```{r}

area_pos_yes <- which(dataset$AREA.1920s != "")
area_pos_no <- which(is.na(dataset$AREA.1920s) == TRUE)

dataset$area_flag <- c(rep(NA, nrow(dataset)))
dataset[area_pos_yes, "area_flag"] <- "yes"
dataset[area_pos_no, "area_flag"] <- "no"

area_plot3 <- dataset %>%
  group_by(UD.1920s, area_flag) %>%
  summarize(count = n())

test <- dataset[dataset$AREA.1920s == "NA",]

totals<- area_plot3 %>%
	group_by(UD.1920s) %>%
	summarise(total=sum(count))

# Stacked barplot
area_plot <- ggplot(area_plot3, aes(fill=area_flag, y=count, x=UD.1920s)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual(name = "Area\ndefined", labels = c("no", "yes"), values=c('#999999','darkblue')) +
    geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150, size = 3) +
    labs(title = "AREA.1920s", x="Urban district (UD.1920s)", y="Number of data entries")+
    scale_x_continuous(breaks = c(1:21))+
  theme(plot.title = element_text(face = "bold"))
area_plot 


```

## POS.1920s

```{r}


pos_plot_data <- dataset %>%
  group_by(UD.1920s, POS.1920s) %>%
  summarize(count = n())

pos_plot_data <- data.frame(pos_plot_data, stringsAsFactors = F)

totals<- pos_plot_data  %>%
	group_by(UD.1920s) %>%
	summarise(total=sum(count))

position_pos <- levels(as.factor(pos_plot_data$POS.1920s))

pos_plot <- ggplot(pos_plot_data, aes(fill=factor(POS.1920s, levels = position_pos[c(1,3,5,6,2,4)]), y=count, x=UD.1920s)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual(name = "Data pattern", labels = c("Data not\navailable", position_pos[c(3,5,6,2,4)]), values = c('#999999','#F0F8FF','#E6E6FA','#87CEFA','#483D8B','darkblue')) +
    geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150, size = 3) +
    labs(title = "POS.1920s", x="Urban district (UD.1920s)", y="Number of data entries")+
    scale_x_continuous(breaks = c(1:21))+
  theme(plot.title = element_text(face = "bold"))
pos_plot

```

## FLOORS.1920s
```{r}


floors_plot_bez <- dataset %>%
  group_by(UD.1920s, FLOORS.1920s) %>%
  summarize(count = n())

floors_plot_bez <- data.frame(floors_plot_bez, stringsAsFactors = F)


totals<- floors_plot_bez  %>%
	group_by(UD.1920s) %>%
	summarise(total=sum(count))

floor_plot <- ggplot(floors_plot_bez, aes(fill=FLOORS.1920s, y=count, x=UD.1920s)) + 
    geom_bar(position="stack", stat="identity") +
    geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150, size = 3) + 
    labs(title = "FLOORS.1920s", x="Urban district (UD.1920s)", y="Number of data entries", fill ="Number of floors\nabove ground floor")  +
    #scale_fill_manual(name = "Data class")+  
     scale_x_continuous(breaks = c(1:21))+
    theme(plot.title = element_text(face = "bold")) +
    geom_point(aes(x = 1, y = 1, size = "Data not\navailable"), shape = NA, colour = "grey") +
    guides(size = guide_legend("", override.aes = list(shape = 15, size = 7)))
floor_plot

```

## YoC.1920s

```{r}

# Create categories

yoc_salz_pos_1 <- grep("^\\d{4}$", dataset$YoC.1920s) # nur 4 stellige Zahlen
yoc_salz_pos_2 <- grep("^\\d{4}[,]", dataset$YoC.1920s) # nur 4 stellige Zahlen am Beginn + ein ,
yoc_salz_pos_3 <- which(dataset$YoC.1920s == "")

dataset$yoc_plot_bez <- rep(NA, nrow(dataset))
dataset[yoc_salz_pos_1, "yoc_plot_bez"] <- "One year date"
dataset[yoc_salz_pos_2, "yoc_plot_bez"] <- "Two year date"
dataset[yoc_salz_pos_3, "yoc_plot_bez"] <- "not available"

# Create data

yoc_plot_bez <- dataset %>%
  group_by(UD.1920s, yoc_plot_bez) %>%
  summarize(count = n())

yoc_plot_bez <- data.frame(yoc_plot_bez, stringsAsFactors = F)

totals<- yoc_plot_bez  %>%
	group_by(UD.1920s) %>%
	summarise(total=sum(count))

yoc_plot_bez_fig_factor <- levels(as.factor(yoc_plot_bez$yoc_plot_bez))

yoc_plot <- ggplot(yoc_plot_bez, aes(fill= factor(yoc_plot_bez, levels = yoc_plot_bez_fig_factor[c(1,3,2)]), y=count, x=UD.1920s)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual(name = "Data pattern", values=c('#999999','lightblue','darkblue')) + 
    geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150 , size = 3) +
    labs(title = "YoC.1920s", x="Urban district (UD.1920s)", y="Number of data entries") +
    scale_x_continuous(breaks = c(1:21))+
    theme(plot.title = element_text(face = "bold"))
yoc_plot

```

## YoP.1920s
```{r}

# Create categories
yop_salz_pos_1 <- grep("^\\d{4}$", dataset$YoP.1920s) # nur 4 stellige Zahlen
yop_salz_pos_2 <- grep("^\\d{4}[,]", dataset$YoP.1920s) # nur 4 stellige Zahlen am Beginn + ein ,
yop_salz_pos_3 <- which(dataset$YoP.1920s == "")

dataset$yop_plot_bez <- rep(NA, nrow(dataset))
dataset[yop_salz_pos_1, "yop_plot_bez"] <- "One year date"
dataset[yop_salz_pos_2, "yop_plot_bez"] <- "Two year date"
dataset[yop_salz_pos_3, "yop_plot_bez"] <- "not available"

# Create data
yop_plot_bez_fig <- dataset %>%
  group_by(UD.1920s, yop_plot_bez) %>%
  summarize(count = n())

yop_plot_bez_fig <- data.frame(yop_plot_bez_fig, stringsAsFactors = F)

totals<- yop_plot_bez_fig  %>%
	group_by(UD.1920s) %>%
	summarise(total=sum(count))

yop_plot_bez_fig_factor <- levels(as.factor(yop_plot_bez_fig$yop_plot_bez))

yop_plot <- ggplot(yop_plot_bez_fig, aes(fill = factor(yop_plot_bez, levels = yop_plot_bez_fig_factor[c(1,3,2)]), y=count, x=UD.1920s)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual(name = "Data pattern", values=c('#999999','lightblue','darkblue')) +
    geom_text(data=totals, aes(x=UD.1920s, label=total, y=total, fill=NULL), nudge_y=150, size = 3) +
    labs(title="YoP.1920s", x="Urban district (UD.1920s)", y="Number of data entries") +
    scale_x_continuous(breaks = c(1:21))+
  theme(plot.title = element_text(face = "bold"))
yop_plot

```
## PDF.pages
```{r}


cadastral$PDF.page.nr <- cadastral$PDF.page.end - cadastral$PDF.page.start + 1

# Page count: Anaolog building schematic
kg1 <- aggregate(PDF.page.nr ~ UD.1920s, cadastral, sum)

kg12 <- sum(kg1$PDF.page.nr, na.rm = TRUE) # total page number

cadastral$vol_par <- paste(as.character(cadastral$UD.1920s), as.character(cadastral$Volume), as.character(cadastral$Part), sep = "-")

###############

pdf <- unique(dataset[,c("UD.1920s", "Page.pdf")])
pdf <- pdf[order(pdf$UD.1920s, pdf$Page.pdf),]
pdf$Page.pdf <- as.integer(pdf$Page.pdf)


# add volume
volume_raw <- unique(cadastral[,c("Volume", "UD.1920s")])
volume_raw$Volume <- as.integer(volume_raw$Volume)
volume_raw$UD.1920s <- as.integer(volume_raw$UD.1920s)
pdf <- merge(pdf,volume_raw, by.x = "UD.1920s", by.y = "UD.1920s", sort = T)

# Page number per urban district
totals<- kg1[,c("UD.1920s", "PDF.page.nr")]

cadastral$PDF.page.end <- as.integer(cadastral$PDF.page.end)
cadastral$UD.1920s <- as.integer(cadastral$UD.1920s)

max_df <- cadastral %>%
	group_by(UD.1920s) %>%
	summarise(max=max(PDF.page.end))
max_df <- data.frame(max_df, stringsAsFactors = F)
max_df$UD.1920s <- as.integer(max_df$UD.1920s)
max_df$max <- as.integer(max_df$max)
max_df <- max_df[order(max_df$UD.1920s),]

totals$max <- max_df$max
totals$max <- as.numeric(totals$max)
totals$label_text <- paste(rep("[",21),totals$PDF.page.nr, "]",sep = "")

pdf_plot <-ggplot(data=pdf, aes(x=UD.1920s, y=Page.pdf)) +
  geom_point(aes(colour = factor(Volume)), size = 0.1, shape=0) +
  labs(title = "Page.pdf", x="Urban district (UD.1920s)", y="Page number (Page.pdf)\n[page count]") +
  scale_colour_discrete("Volume of analog\nbuilding schematic") +
  guides(color = guide_legend(override.aes = list(size=3, shape = rep(15,10))))+
  theme(plot.title = element_text(face = "bold", family = "Helvetica"))+
  scale_x_continuous(breaks = c(1:21))+
  geom_text(data=totals, aes(x=UD.1920s, label= label_text, y = max), nudge_y=10, size =3)
pdf_plot

```


# External validation
## Data Completness: Number of buildings by urban district
This code section produces Figure 7 in the Data Descriptor.

```{r}

a <- which(dataset$BN.1920s != "") # Einträge mit BN.1920s
b <- which(dataset$BN.1920s == "")

c <- dataset[a,] %>%
  group_by(UD.1920s) %>%
  summarize(n())
c <- data.frame(c)
colnames(c) <- c("UD.1920s", "counts")
c1 <- aggregate(counts ~ UD.1920s, c, sum)

cd2 <- cbind.data.frame(c1, yearbook_1923$Häuser_2) # Es werden die Daten vom Statischtischen Jahrbuch genommen.https://www.digital.wienbibliothek.at/wbrobv/periodical/pageview/2176992
cd2$diff.abs <- cd2[,2]-cd2[,3]
cd2$diff.rel <- round(cd2[,4]/cd2[,3],2)
colnames(cd2) <- c("UD.1920s", "counts.salzberg", "counts.stat", "diff.abs", "diff.rel")

cd2.sum <- data.frame(sum(cd2$counts.salzberg), sum(cd2$counts.stat), sum(cd2$diff.abs), (sum(cd2$diff.abs) / sum(cd2$counts.stat)))

# Plotting figure

data.salz <- data.frame(c(1:21), rep("building schematic", 21), cd2[,2])
data.stat <- data.frame(c(1:21), rep("census", 21), cd2[,3])
coln <- c("UD.1920s", "data_source", "counts")
colnames(data.salz) <- coln
colnames(data.stat) <- coln

data <- rbind(data.stat, data.salz)

p <- ggplot(data, aes(fill=data_source, y=counts, x=UD.1920s)) + 
    geom_bar(position="dodge", stat="identity") +
    ggtitle("Comparative building counts") + 
    scale_x_continuous(labels = c(1:21), breaks = c(1:21)) +
    labs(y= "Number of buildings", x = "Urban district") +
    scale_fill_discrete(name = "Data source", labels = c("Statistical yearbook (1923)", "Digital building schematic (1927-30)")) + 
    theme(legend.position = c(0.23,0.85))
p

```

## Data plausibility: Number of buildings by floor counts
This code section produces Figure 8 in the Data Descriptor.

```{r}

floors_salz <- dataset %>%
  group_by(FLOORS.1920s) %>%
  summarise(anz = n())

floors_salz$cum <- cumsum(floors_salz$anz)

# Validierung mit Stockwerksstatitisk von 1914

# Gebäude filtern (Integer, Integer and letter, letter and integer)
buildings_w_bn_pos <- which(dataset$bn.table.ud != "Data not available")

floors <- dataset[buildings_w_bn_pos,] %>%
  group_by(FLOORS.1920s) %>%
  summarize(n_salz = n())

floors$FLOORS.1920s <- as.character(floors$FLOORS.1920s)

floors[7,1] <- "unknown" # Mache NA zu "unknown""

# Gebäude nach Stockwerken 1914 einlesen

floors_1914_wien <- data.frame(colSums(floors_1914[,2:8]))

floors_complete_1914 <- cbind(c("ground floor only", "1", "2", "3", "4", "5 or more", "unknown"), rep("Statistical yearbook (1914)",7), floors_1914_wien$colSums.floors_1914...2.8..)
colnames(floors_complete_1914) <- c("Floors", "data_source", "count")                                                                                          

floors_complete_1920s <- cbind(c("ground floor only", "1", "2", "3", "4", "5 or more", "unknown"), rep("Digital building schematic (1927-30)",7), c(NA, floors[1,2],floors[2,2],floors[3,2],floors[4,2], sum(floors[5:6,2]), floors[7,2]))


colnames(floors_complete_1920s) <- c("Floors","data_source", "count")             
                                                                                                                      
floors_complete <- data.frame(rbind(floors_complete_1914, floors_complete_1920s), stringsAsFactors = F)

rownames(floors_complete) <- NULL
floors_complete$count <- as.numeric(floors_complete$count)
floors_complete$Floors <- as.character(floors_complete$Floors)

positions <- c("ground floor only", "1", "2", "3", "4", "5 or more", "unknown")
floors_complete$count_label <- floors_complete$count
floors_complete$count_label <- as.character(floors_complete$count_label)

floors_complete[8,3] <- 0
floors_complete[8,4] <- "NA"

# plot
floor_valid_plot <- ggplot(floors_complete, aes(x=Floors, y=count, fill=factor(data_source, levels = c("Statistical yearbook (1914)", "Digital building schematic (1927-30)"))))+
  geom_bar(position="dodge", stat="identity")+
  scale_x_discrete(limits = positions)+  
  ggtitle("Comparative building counts") + 
  labs(y= "Number of buildings", x = "Floors") +
  scale_fill_discrete(name = "Data source", labels = c("Statistical yearbook (1914)", "Digital building schematic (1927-30)")) +   
  theme(legend.position = c(0.23,0.85))+
  coord_cartesian(ylim = c(0, 15000))+
  geom_text(aes(label=count_label),  position = position_dodge(0.9), vjust=-1, size = 3)
floor_valid_plot

```
